---
title: "Do Federal Reserve Banks Research What Their Districts Produce?"
subtitle: "Comparing Fed working paper topics to district industry composition"
author: "Benedict Leonardi"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    fig-width: 9
    fig-height: 6
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(ggrepel)

topic_shares    <- read_csv("data/fed_topic_shares.csv")
industry_shares <- read_csv("data/fed_industry_shares.csv")
pair_cors       <- read_csv("data/pair_correlations.csv")
cross_cor       <- read_csv("data/cross_correlation_matrix.csv")
alignment       <- read_csv("data/fed_alignment_scores.csv")
joined          <- read_csv("data/joined_wide.csv")

sectors <- c("Housing","Energy","Agriculture","Technology","Finance","Trade","Manufacturing","Labor")

theme_set(
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40"),
    panel.grid.minor = element_blank()
  )
)
```

## Overview

This analysis asks whether Federal Reserve Banks' research output mirrors the
economic specialization of their geographic districts. Each Fed publishes working
papers through its research department; if researchers respond to their local
economy, we'd expect — for example — the Dallas Fed to publish more energy
research, or the New York Fed to publish more finance research.

**Data sources:**

- **Working papers (2010–2024):** OpenAlex API, ~18,300 papers across all 12 Feds
- **Industry employment:** BLS QCEW 2022 annual county-level data, aggregated to
  Fed districts using official district-county crosswalks
- **Topic model:** keyATM with 10 seeded topics (Agriculture, Energy, Finance,
  Housing, Labor, Manufacturing, MonetaryPolicy, Technology, Trade, Inequality)
  plus 3 unsupervised residual topics

---

## Topic mix by Fed

How does each Fed's research portfolio break down by topic?

```{r topic-heatmap, fig.height=7}
topic_long <- topic_shares |>
  select(fed, all_of(sectors)) |>
  pivot_longer(-fed, names_to = "topic", values_to = "share")

# Normalize within topic (z-score) so differences are visible across topics
topic_long <- topic_long |>
  group_by(topic) |>
  mutate(z = (share - mean(share)) / sd(share)) |>
  ungroup()

ggplot(topic_long, aes(x = topic, y = fct_reorder(fed, share, .fun = sum), fill = z)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = scales::percent(share, accuracy = 0.1)), size = 2.8, color = "black") +
  scale_fill_gradient2(
    low = "#3B82F6", mid = "white", high = "#EF4444",
    midpoint = 0, name = "Z-score\n(within topic)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 35)) +
  labs(
    title    = "Research topic mix by Federal Reserve Bank",
    subtitle = "Cell values = share of papers; color = deviation from cross-Fed mean within topic",
    x = NULL, y = NULL
  )
```

---

## Industry mix by Fed district

How does employment break down across the same broad sectors?

```{r industry-heatmap, fig.height=7}
ind_long <- industry_shares |>
  filter(sector %in% sectors)

ind_long <- ind_long |>
  group_by(sector) |>
  mutate(z = (emp_share - mean(emp_share)) / sd(emp_share)) |>
  ungroup()

ggplot(ind_long, aes(x = sector, y = fct_reorder(fed, emp_share, .fun = sum), fill = z)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = scales::percent(emp_share, accuracy = 0.1)), size = 2.8, color = "black") +
  scale_fill_gradient2(
    low = "#3B82F6", mid = "white", high = "#EF4444",
    midpoint = 0, name = "Z-score\n(within sector)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 35)) +
  labs(
    title    = "Industry employment mix by Federal Reserve district",
    subtitle = "Cell values = share of district employment; color = deviation from cross-Fed mean",
    x = NULL, y = NULL
  )
```

---

## Matched-pair correlations

For each topic-industry pair, how strongly does the cross-Fed ranking of
research emphasis correlate with the cross-Fed ranking of industry employment?

```{r pair-cors, fig.height=5}
pair_cors |>
  mutate(
    sig   = p < 0.1,
    label = paste0("r = ", round(r, 2), if_else(sig, "*", ""))
  ) |>
  ggplot(aes(x = r, y = fct_reorder(sector, r), fill = r > 0)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = label, hjust = if_else(r > 0, -0.1, 1.1)), size = 3.5) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  scale_fill_manual(values = c("#3B82F6","#EF4444"), guide = "none") +
  scale_x_continuous(limits = c(-0.65, 0.65)) +
  labs(
    title    = "Cross-Fed correlation: research topic share vs industry employment share",
    subtitle = "n = 12 Feds; * p < 0.1",
    x = "Pearson r", y = NULL
  )
```

**Key findings:**

- **Finance** has the strongest positive match (r = `r round(filter(pair_cors, sector=="Finance")$r, 2)`):
  Feds in finance-heavy districts (New York, Philadelphia, Richmond) publish
  proportionally more finance research.
- **Trade** is also positive (r = `r round(filter(pair_cors, sector=="Trade")$r, 2)`):
  St. Louis and Minneapolis — major logistics/trade corridors — publish more
  trade research.
- **Housing** is *negative* (r = `r round(filter(pair_cors, sector=="Housing")$r, 2)`):
  The Feds with the most housing-sector employment are *not* the ones publishing
  the most housing research. Philadelphia publishes the most housing research
  despite being a mid-tier construction market.
- **Energy** is near zero (r = `r round(filter(pair_cors, sector=="Energy")$r, 2)`):
  Dallas leads on energy employment but San Francisco leads on "energy" research
  topics — likely because San Francisco's energy keywords capture climate/grid
  research rather than oil & gas.

---

## Scatter plots: topic vs industry for key pairs

```{r scatter-pairs, fig.height=8, fig.width=10}
plot_scatter <- function(sector_name, color = "#3B82F6") {
  ggplot(joined, aes(
    x = .data[[paste0(sector_name, "_ind")]],
    y = .data[[paste0(sector_name, "_topic")]]
  )) +
    geom_smooth(method = "lm", se = TRUE, color = "grey70", fill = "grey90", linewidth = 0.7) +
    geom_point(size = 3, color = color, alpha = 0.85) +
    geom_text_repel(aes(label = fed), size = 3, max.overlaps = 20) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
    labs(
      title = sector_name,
      x = "Industry employment share",
      y = "Research topic share"
    )
}

library(patchwork)
(plot_scatter("Finance", "#3B82F6") | plot_scatter("Energy",  "#F59E0B")) /
(plot_scatter("Housing", "#EF4444") | plot_scatter("Trade",   "#10B981"))
```

---

## Full cross-correlation matrix

Do topics correlate with *unexpected* industries? The off-diagonal cells are
often as interesting as the diagonal.

```{r cross-cor-heatmap, fig.height=7}
cross_cor |>
  ggplot(aes(x = sector, y = fct_rev(topic), fill = r)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = round(r, 2)), size = 3) +
  scale_fill_gradient2(
    low = "#3B82F6", mid = "white", high = "#EF4444",
    midpoint = 0, limits = c(-1, 1), name = "r"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 35)) +
  labs(
    title    = "Cross-correlation: every research topic vs every industry",
    subtitle = "Diagonal = matched pairs; off-diagonal = unexpected correlations",
    x = "Industry employment", y = "Research topic"
  ) +
  # highlight diagonal
  geom_tile(
    data = cross_cor |> filter(topic == sector),
    color = "black", fill = NA, linewidth = 1
  )
```

**Notable off-diagonal findings:**

- **Energy research ~ Agriculture employment (r = 0.92):** The Feds in
  agricultural districts (Minneapolis, Kansas City) publish the most
  "energy-tagged" research. This likely reflects rural energy topics —
  ethanol, farm energy costs, rural utilities — rather than oil & gas.
- **Finance research ~ Labor employment (r = 0.40):** Feds in large labor
  markets publish more finance research, possibly reflecting broader
  economic size.
- **Housing research ~ Finance employment (r = 0.37):** Finance-heavy
  districts study housing markets (mortgages, housing finance).

---

## Per-Fed alignment scores

How correlated is each Fed's *overall* topic distribution with its
*overall* industry mix?

```{r alignment, fig.height=5}
alignment |>
  mutate(
    direction = if_else(alignment_score > 0, "Aligned", "Misaligned"),
    label     = round(alignment_score, 2)
  ) |>
  ggplot(aes(
    x = alignment_score,
    y = fct_reorder(fed, alignment_score),
    fill = direction
  )) +
  geom_col(width = 0.6) +
  geom_text(aes(label = label, hjust = if_else(alignment_score > 0, -0.1, 1.1)), size = 3.5) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  scale_fill_manual(values = c("Aligned" = "#10B981", "Misaligned" = "#EF4444"), guide = "none") +
  scale_x_continuous(limits = c(-0.35, 0.6)) +
  labs(
    title    = "Per-Fed alignment: research topic mix vs district industry mix",
    subtitle = "Score = Pearson r between each Fed's topic vector and its industry employment vector",
    x = "Alignment score (r)", y = NULL
  )
```

**Interpretation:**

- **Midwest Feds (St. Louis, Kansas City, Minneapolis)** are most aligned —
  their research portfolios broadly match their districts' manufacturing,
  agriculture, and trade emphasis.
- **Boston, Philadelphia, San Francisco** are most misaligned — their research
  diverges from local industry. Philadelphia over-indexes on housing and health
  research; San Francisco over-indexes on monetary policy relative to its
  tech-heavy economy; Boston research is broadly national in scope.
- **New York** is near zero — it's the most nationally-focused Fed, so neither
  aligned nor misaligned with any specific sector.

---

## Caveats

1. **Abstract coverage:** Only 59% of papers had abstracts in OpenAlex. Papers
   without abstracts were classified on title alone, which is noisier.
2. **"Energy" topic contamination:** San Francisco's high Energy topic score likely
   reflects climate/electricity research, not fossil fuels. A finer keyword set
   (separating fossil-fuel energy from clean energy) would improve this.
3. **State-level district assignment:** Split states (Missouri, West Virginia, etc.)
   were assigned at the county level using published Fed district maps, but a few
   edge counties may be misassigned.
4. **No causal claim:** Correlation between research topics and local industries
   could run in either direction, or both could reflect third factors (e.g.,
   Fed economists' research interests attracted by local data availability).

